---
title: "Investigate the exponential distribution in R and compare it with the CLT"
author: "Philip Wong"
date: "29/05/2022"
output: pdf_document
---

## Overview
In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials. Note that it will require 1000 simulations.

## Simulations
Set the simulation variables lambda, exponentials, and seed.
```{r}
ECHO=TRUE
set.seed(1337)
lambda = 0.2
exponentials = 40
```

Run Simulations with variables
```{r}
simulation_Means = NULL
for (i in 1 : 1000) simulation_Means = c(simulation_Means, mean(rexp(exponentials, lambda)))
```

## Sample Mean versus Theoretical Mean
#### Sample Mean
```{r}
mean(simulation_Means)
```

#### Theoretical Mean
```{r}
lambda^-1
```

#### Comparison
```{r}
abs(mean(simulation_Means)-lambda^-1)
```

## Sample Variance versus Theoretical Variance
#### Sample Variance
```{r}
var(simulation_Means)
```

#### Theoretical Variance
```{r}
(lambda * sqrt(exponentials))^-2
```

#### Comparison
```{r}
abs(var(simulation_Means)-(lambda * sqrt(exponentials))^-2)
```


## Distribution
```{r}
library(ggplot2)
ggplot(data.frame(y=simulation_Means), aes(x=y)) + 
  geom_histogram(aes(y=..density..), binwidth=0.2, fill="dark green", 
                 color="black") +
  stat_function(fun=dnorm, arg=list(mean=lambda^-1, 
                                    sd=(lambda*sqrt(exponentials))^-1), 
                size=2) +
  labs(
    title = "Plot of the Simulations", 
    x = "Simulation Mean")
```